---
title: "To Generate a **.csv** file from a chess tournament results"
author: "Nnaemeka Okereafor"
date: "9/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 
### About the project
In this project, we're given a text file with chess tournament results where the information has some structure. Our job was to create an R Markdown file that generates a *.CSV* file (that could for example be imported into a SQL database) with the following information for all of the players:
**Player’s Name**, **Player’s State**, **Total Number of Points**, **Player’s Pre-Rating**, and **Average Pre Chess Rating of Opponents**. Illustration, for the first player, the information would be: Gary Hua, ON, 6.0, 1794, 1605

```{r warning=FALSE, message=FALSE}
library(readr)
library(stringr)

df1<-read.table(file="https://raw.githubusercontent.com/nnaemeka-git/global-datasets/main/tournamentinfo.txt",stringsAsFactors = FALSE, sep=",",skip=4)
```

### To remove **dashes**

```{r}
df2 <- df1$V1
df2 <- str_replace_all(df2, pattern = "\\s+", replacement = " " )
df2 <- str_remove_all(df2, pattern = "-")
```


### Names

To match **Players Names**

```{r  warning=FALSE, message=FALSE}
names <- unlist(str_match_all(df1, " [A-Za-z]{1,}\\s*[A-Za-z.,-]{1,}\\s*[A-Za-z.,-]{1,}\\s?"))
names

```

### States

To match the names of the **States**

```{r  warning=FALSE, message=FALSE}
state <- unlist(str_match_all(df1, " [A-Za-z]{2}\\s\\|"))
state <-str_remove_all(state,"\\|")
state


```

### Points

To match the **Player Points**

```{r  warning=FALSE, message=FALSE}
pts <- unlist(str_match_all(df1,"\\|\\d+.."))
#To remove pipe
pts <-str_remove_all(pts,"\\|")
pts

```


### Pre Rating

To match the **pre-rating scores**

```{r  warning=FALSE, message=FALSE}
Pre_rate <-unlist(str_match_all(df1,"R.\\s.\\d+"))
Pre_rate <-as.numeric(str_remove_all(Pre_rate,'R:'))

# Convert into a dataframe
Pre_rate<-data.frame(Pre_rate)
colnames(Pre_rate)<-"rating"
index<-data.frame(as.numeric(row.names(Pre_rate)))
colnames(index)<-"Index"
pre_rate_df <- data.frame(index,Pre_rate)

head(pre_rate_df)

```

### Number of Games Played

To extract the **number of games played and the opponents' numbers**

```{r  warning=FALSE, message=FALSE}
num_game <- unlist(str_extract_all(df2,"\\|[0-9].*"))
num_game <- unlist(str_replace_all(num_game, "\\|[BUXH] ", replacement = "\\|R 0"))
```

```{r}
#To remove the first part containing the total points
oppon <- unlist(str_remove_all(num_game, pattern = "\\|\\d\\.\\d\\s"))

#To remove the alphabets and the pipe
oppone <- unlist(str_remove_all(oppon, pattern = "[:alpha:]\\|"))

oppone <- as.numeric(unlist(str_extract_all(oppone, pattern = "[:digit:]{1,2}")))
head(oppone)
```

### To convert vectors into matrix and add columns to the matrix to represent the Opponents in each round of game

```{r}
col_names=c("O1","O2","O3","O4","O5","O6","O7")
op_matr <- as.data.frame(matrix(oppone,byrow=TRUE,ncol=7,))
colnames(op_matr)<-col_names
op_matr

```

###  To extract the **opponents's pre-rating scores**

```{r}
#nexted forloop does it better

for (row in 1:nrow(op_matr)){
  for (col in 1:ncol(op_matr)){
    if (op_matr[row,col] != 0){
      op_matr[row,col] = pre_rate_df$rating[op_matr[row,col]]
    } else {
       op_matr[row,col] = NA
     }
  }
}

op_matr
```

### To label **opponents' pre-rating scores** columns, count **Number of Opponents**, sum the **total opponents' pre-rating score** and calculate the **Average opponents' pre-rating score**

```{r}
matr_names=c("O1 Pre-rating","O2 Pre-rating","O3 Pre-rating","O4 Pre-rating","O5 Pre-rating","O6 Pre-rating","O7 Pre-rating")
colnames(op_matr)<-matr_names
op_matr$Num.of.Opponents <- rowSums(!is.na(op_matr))
op_matr$TotalPreOp_rating <- rowSums(op_matr[,1:7],na.rm=TRUE)
op_matr$AvgPreOp_rating <- round((op_matr$TotalPreOp_rating/op_matr$Num.of.Opponents),0)
head(op_matr,n=20)
```

### Combine dataframes to produce a single table

```{r  warning=FALSE, message=FALSE}
library(dplyr)
table <- data.frame("Player Name"=names, state=state, "Number of Points"=pts, pre_rate_df,op_matr)
head(table,n=15)
```

### Create a table with the **required columns** from the final table

```{r}
chess_rating_table <- table%>% select(Player.Name,state,Number.of.Points,rating,AvgPreOp_rating)
chess_rating_table
```

### Write the dataframe to a **csv** file

```{r}
write.csv(chess_rating_table, row.names = FALSE)

```

### Writing **chess_rating_table** into a directory

```{r}
#write.csv(chess_rating_table, "choice_file_name.csv", row.names = FALSE)
```

